---
title: "J. Alex Baecher"
subtitle: "Curriculum Vitae"
output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: false
    includes:
      in_header: cv-header.tex
  html_document:
    theme: cosmo
    css: cv-style.css
fontsize: 11pt
geometry: margin=0.75in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)
```

\begin{center}
\faEnvelope\ jbaecher@gmail.com | \faPhone\ 352-226-3983 | \faGlobe\ \href{https://alexbaecher.com}{alexbaecher.com}
\end{center}

# Education

**Doctor of Philosophy, Interdisciplinary Ecology** \hfill Aug 2024  
*University of Florida*

**Master of Science, Biology** \hfill Dec 2017  
*Eastern Kentucky University*

**Bachelor of Science, Biology** \hfill May 2014  
*University of Arkansas*

# Experience

**Postdoctoral Fellow in Quantitative Ecology** \hfill July 2025 – Present  
*Utah State University*

**Postdoctoral Research Associate** \hfill Aug 2024 – July 2025  
*University of Florida*

**Doctoral Preeminence Fellow** \hfill Aug 2019 – 2024  
*University of Florida*

**Conservation and Research Technician** \hfill Jan 2018 – July 2019  
*Memphis Zoo*

**NSF REU Research Coordinator** \hfill May – Aug 2017  
*Eastern Kentucky University*

**Graduate Teaching Assistant** \hfill May 2015 – Aug 2017  
*Eastern Kentucky University*

**Undergraduate Research Assistant** \hfill Aug 2012 – Mar 2015  
*University of Arkansas*

# Funding

```{r funding-table}
funding <- tribble(
  ~Amount, ~Organization, ~Fund, ~Year,
  "$130,000", "US Geological Survey", "Climate Adaptation Postdoctoral Fellows", "2025",
  "$195,000", "University of Florida", "Doctoral Preeminence Fellowship Award", "2019",
  "$4,000", "University of Florida", "Robin Nadeau Graduate Research Award", "2019",
  "$5,000", "National Science Foundation", "Research Experience for Undergraduates", "2013",
  "$5,000", "National Science Foundation", "Research Experience for Undergraduates", "2012",
  "$1,100", "Eastern Kentucky University", "Marcia Athey Research Grant", "2016",
  "$800", "Eastern Kentucky University", "Dept. Biology Travel Grant", "2017",
  "$700", "Society Freshwater Science", "Instars Fellowship", "2014",
  "$700", "Society Freshwater Science", "Instars Fellowship", "2013",
  "$500", "Soc. Study Amphibians & Reptiles", "Grant-in-Aid of Herpetological Research", "2016"
)

if(knitr::is_latex_output()) {
  kable(funding, 
        format = "latex",
        booktabs = TRUE,
        align = c("r", "l", "l", "r"),
        col.names = c("Amount", "Organization", "Fund Name", "Year")) %>%
    kable_styling(latex_options = c("scale_down", "HOLD_position"),
                  font_size = 10)
} else {
  kable(funding,
        format = "html",
        align = c("r", "l", "l", "r"),
        col.names = c("Amount", "Organization", "Fund Name", "Year")) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
}
```

**Total Funding: \$344,300**

# Publications

*§Mentee co-author • Citations: 277 • h-index: 7 • Impact Factor: 8.7 (average)*

## Published

15. **Baecher, JA**, §VA Akshay, RP Guralnick, §AM Bauer, §Y Tavares, §Y Sánchez, LP Campbell. 2025. Toward ecological forecasting of West Nile virus in Florida: Insights from two decades of sentinel chicken monitoring. *Science of the Total Environment* 1000:180308 [\[HTML\]](https://doi.org/10.1016/j.scitotenv.2024.180308) [\[PDF\]](https://doi.org/10.1016/j.scitotenv.2024.180308) [\[CODE\]](https://github.com/)

14. Evans, LJ, **JA Baecher**, BR Scheffers. 2024. Invasion risk posed by the pet trade. *Frontiers in Ecology and the Environment* [\[HTML\]](https://doi.org/) [\[PDF\]](https://doi.org/)

13. **Baecher, JA**. 2024. The environmental filtering paradigm and non-filtering community assembly processes. *Journal of Biogeography* 51(11):2109–2116 [\[HTML\]](https://doi.org/) [\[PDF\]](https://doi.org/)

12. Klinges, DH, **JA Baecher**, JJ Lembrechts, IMD Maclean, J Lenoir, C Greiser, M Ashcroft, LJ Evans, MR Kearney, J Aalto, LC Barrio, P De Frenne, J Guillemot, K Hylander, T Jucker, M Kopecky, M Luoto, M Macek, I Nijs, J Urban, L van den Brink, P Vangansbeke, J Von Oppen, J Wild, J Boike, R Canessa, M Nosetto, A Rubtsov, J Sallo-Bravo, BR Scheffers. 2024. Proximal microclimate: moving beyond spatiotemporal scale to improve predicted biological responses. *Global Ecology and Biogeography* 33(7):e1–16 [\[HTML\]](https://doi.org/) [\[PDF\]](https://doi.org/)

11. Comte, L, R Bertrand, SE Diamond, LT Lancaster, ML Pinsky, BR Scheffers, **JA Baecher**, RMWJ Bandara, I-C Chen, JA Lawlor, NA Moore, BR Oliveira, J Murienne, J Rolland, MA Rubenstein, J Sunday, LM Thompson, F Villalobos, SR Weiskopf, J Lenoir. 2024. Bringing traits back in the equation: A roadmap to understand species redistribution. *Global Change Biology* 30:e17271 [\[HTML\]](https://doi.org/) [\[PDF\]](https://doi.org/)

10. Lawlor, JA, L Comte, G Grenouillet, J Lenoir, **JA Baecher**, RMWJ Bandara, R Bertrand, I-C Chen, SE Diamond, LT Lancaster, NA Moore, J Murienne, BF Oliveria, GT Pecl, ML Pinsky, J Rolland, MA Rubenstein, BR Scheffers, LM Thompson, B can Amerom, F Villalobos, SR Weiskopf, J Sunday. 2024. Mechanisms, detections, and impacts of species redistributions under climate change. *Nature Reviews Earth & Environment* 5:351–368 [\[HTML\]](https://doi.org/) [\[PDF\]](https://doi.org/) | *Impact Factor: 71.5*

9. **Baecher, JA**, SA Johnson, EA Roznik, BR Scheffers. 2023. Experimental evaluation of how invasions and climate change interact to alter the vertical assembly of an amphibian community. *Journal of Animal Ecology* 92(4):875–888 [\[HTML\]](https://doi.org/) [\[PDF\]](https://doi.org/) [\[CODE\]](https://github.com/)

8. Basham, EW, **JA Baecher**, DH Klinges, BR Scheffers. 2022. Vertical stratification patterns of tropical forest vertebrates: A meta-analysis. *Biological Reviews* 98(1):99–114 [\[HTML\]](https://doi.org/) [\[PDF\]](https://doi.org/)

7. Basham, EW, M González-Pinzón, A Romero-Marcucci, N Carl, **JA Baecher**, BR Scheffers. 2022. Large, old trees define the vertical, horizontal, and seasonal distributions of a poison frog. *Oecologia* 199:257–269 [\[HTML\]](https://doi.org/) [\[PDF\]](https://doi.org/)

6. Swartwout, M, PN Vogrinc, **JA Baecher**, CS Kross, JD Willson. 2020. Prey Size and Growth Rate do not Influence Trophic Morphology of Juvenile Water Snakes *Nerodia sipedon*. *Herpetologica* 76(1):53–60 [\[HTML\]](https://doi.org/) [\[PDF\]](https://doi.org/)

5. **Baecher, JA**, §E Jones, SC Richter. 2020. Tail bifurcation in two species of *Desmognathus* salamander Caudata: Plethodontidae in southeastern Kentucky, USA. *Herpetological Bulletin* 150:29–30 [\[HTML\]](https://doi.org/) [\[PDF\]](https://doi.org/)]

4. Hinkson, KM, **JA Baecher**, S Poo. 2019. Cryopreservation and hormonal induction of spermic urine in the smooth-sided toad *Rhaebo guttatus*. *Cryobiology* 89:109–111 [\[HTML\]](https://doi.org/) [\[PDF\]](https://doi.org/)]

3. **Baecher, JA**, SC Richter. 2018. Environmental gradients in an old-growth Appalachian Forest predict fine-scale distribution, co-occurrence, and abundance of woodland salamanders. *Ecology and Evolution* 8(24):12940–12952 [\[HTML\]](https://doi.org/) [\[PDF\]](https://doi.org/)]

2. **Baecher, JA**, PH Vogrinc, JC Guzy, CS Kross, JD Willson. 2017. Herpetofaunal Communities in Restored and Unrestored Remnant Tallgrass Prairie and Associated Wetlands in Northwest Arkansas. *Wetlands* 38:157–168 [\[HTML\]](https://doi.org/) [\[PDF\]](https://doi.org/)]

1. **Baecher, JA**, PH Vogrinc, JC Guzy, JC Neal, JD Willson. 2014. *Lithobates areolatus*, Predation. *Herpetological Review* 45(4):681–682 [\[PDF\]](https://doi.org/)]

## In Press & Under Review

- Oliveira, BF, et al. [+21 co-authors including **JA Baecher**]. In press. Species range shifts often speed ahead of their climatic niches. *Proceedings of the National Academy of Sciences*

- **Baecher, JA**, D Klinges, LJ Evans, CA Romagosa, R Fletcher, BR Scheffers. In revision. Isolating what limits the spread of an invasive predator. *Journal of Applied Ecology* [\[CODE\]](https://github.com/) [\[PREPRINT\]](https://doi.org/)

- §Akshay, VA, **JA Baecher**, N Burkett-Cadena, J Thorson, §Y Sánchez, §Y Tavares, §A Bauer, RP Guralnick, LP Campbell. In review. Linking hosts, landscapes, and climate to advance zoonotic arbovirus forecasting. *Nature Communications* [\[CODE\]](https://github.com/) [\[PREPRINT\]](https://doi.org/)]

- Thompson, LM, et al. [+21 co-authors including **JA Baecher**]. In review. The RAD (Resist-Accept-Direct) management of species redistribution under climate change. *Frontiers in Ecology and the Environment*

## In Preparation

- **Baecher, JA**, D Klinges, LJ Evans, BR Scheffers. In prep. Informing species range-shifts with global climate and habitat connectivity. *Prepared for One Earth*

- Klinges, DH, **JA Baecher**, L Soifer, BR Scheffers. In prep. Land modification defines climate change as experienced by life on earth. *Prepared for Nature*

# Professional Activities

## Peer Review

**Total Requested:** 38 | **Total Reviewed:** 25

Biological Conservation, Conservation Science and Practice, Ecography, Ecology, Ecology & Evolution, Frontiers in Ecology and Environment, Herpetological Bulletin, Herpetological Conservation Biology, Herpetological Review, Hydrobiologia, Natural Areas Journal, Nature Ecology & Evolution, Oecologia, Scientific Reports, Wildlife Society Bulletin

## Professional Membership

- Ecological Society of America (2019 – Present)
- International Association of Landscape Ecologists (North America, 2020 – Present)
- Climate Adaptation Science Center (SE & SW, 2020 – Present)
- Southeast RISCC (2022 – Present)
- American Society of Ichthyologists and Herpetologists (2014 – 2021)
- Society for the Study of Amphibians and Reptiles (2014 – 2021)
- Southeastern Partners in Amphibian and Reptile Conservation (2014 – 2021)
- Society of Freshwater Ecology (2012 – 2014)

## Working Groups

- **BioShifts** (2022 – Present; Member, Group Leader)
- **DirtClim** (2021 – Present; Core team)
- **National Synthesis for Ecological Analysis and Synthesis** (2025 – Present; Member, Postdoctoral Fellow)
- **ResearchR** (2025 – Present; President of Scientific Integrity)
- **ZooMu Network** (2024 – Present; Member, Group Leader)
- **USGS Climate Adaptation Science Center** (2025 – 2027; Postdoctoral Fellow)
- Invasive Species Horizon Scan (2019 – 2021; Contributor)

# Teaching & Mentorship

## Instructor of Record

**Environmental Science Lab (EVS 3000L)** \hfill Fall 2022 & 2023  
*University of Florida | School of Natural Resources and Environment*

## Teaching Assistant

**Human Anatomy and Physiology Lab (BIO 307)** \hfill Spring & Fall 2015–2017  
*Eastern Kentucky University | Department of Biology*

## Guest Lectures (Selected)

- **Utah State University** | Landscape Ecology (WILD 6710) \hfill Fall 2025
  - Connectivity algorithms with R
- **University of Florida** | Community Science for Ecologists (WIS 4934/6934) \hfill Spring 2025
  - Spatiotemporal modeling with Bayesian statistics in R
- **University of Florida** | Ecology of Climate Change (WIS 3410) \hfill Spring 2023
  - Achieving landscape connectivity under rapid change
- **University of Florida** | Wildlife of Florida (WIS 3402) \hfill Spring 2022
  - Amphibian diversity & North American salamander diversity and conservation

## Mentorship

**Graduate Students (MS/PhD pursuing):** 14 mentees  
**Undergraduate Students:** 5 mentees  
**Publication Co-authors:** 7 mentees  
**Conference Presenters:** 12 mentees

Selected mentees: V Anand Akshay (PhD), Amy Bauer (PhD), Yasmin Tavares (MS→PhD), Emily Jones (MS), Jocelyn Stalker (MS), Sarah Ashbrook (MS), Peter Grap (MS), Jacob Peters (MS→PhD), Julia Portmann (MS→PhD), Jack Rasmussen (PhD)

# Select Presentations

*†Student mentee presenter • Total presentations: 41*

1. **Baecher, JA**. 2025. Shifting Frontiers: Understanding species abundance range-shifts for informed conservation strategies. Pacific Flyway Partners (hybrid) [\[SLIDES\]](https://)

2. **Baecher, JA**. 2024. Influence of Global Change on Species' Behavior, Movement, and Ranges. SNRE Seminar Series, Gainesville, FL

3. **Baecher, JA**, L Evans, BR Scheffers. 2024. Toward knowledge co-production in invasion modeling: Developing novel interactive expert-elicitation software. Invasive Species Research Institute Conference, Gainesville, FL

4. **Baecher, JA**, D Klinges, L Evans, BR Scheffers. 2023. Informing Species Range-shifts with Global Connectivity. Species On The Move, Ft Myers, FL

5. **Baecher, JA**, L Evans, BR Scheffers. 2022. Predicting the spread of invasive Burmese pythons in Florida. SERISCC Network Webinar (virtual) [\[SLIDES\]](https://) [\[VIDEO\]](https://)

# Volunteer & Outreach

**Total Events:** 26 | **People Reached:** ~1,430 (ages 5 to 60+)

Selected activities: Guest teaching (elementary & middle school), Science fair judging, BioBlitz community science events, Conservation education at Memphis Zoo (~300 reached), Water quality assessments, Career fair outreach

# References

Available upon request

---

*Last updated: `r format(Sys.Date(), "%B %d, %Y")`*
