---
title: "J. Alex Baecher"
subtitle: "Curriculum Vitae"
format:
  pdf:
    documentclass: article
    geometry:
      - margin=0.75in
    fontsize: 11pt
    include-in-header:
      text: |
        \usepackage{enumitem}
        \usepackage{hyperref}
        \usepackage{fontawesome5}
        \setlist[itemize]{leftmargin=*, nosep, label={}}
        \setlength{\parindent}{0pt}
        \setlength{\parskip}{6pt}
        \usepackage{titlesec}
        \titleformat{\section}{\large\bfseries\uppercase}{\thesection}{0em}{}[\titlerule]
        \titleformat{\subsection}{\normalsize\bfseries}{\thesubsection}{0em}{}
        \titlespacing{\section}{0pt}{0pt}{0pt}
        \titlespacing{\subsection}{0pt}{8pt}{4pt}
  html:
    theme: cosmo
    toc: false
    css: cv-style.css
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)
```

::: contact-info
                           \faEnvelope: [jbaecher\@gmail.com](mailto:jbaecher@gmail.com)   \faPhone: (352) 226-3983   \faGlobe: [questlab.dev](https://questlab.dev)
:::

<!-- [\faGithub](https://github.com/slamander) &nbsp;&nbsp;&nbsp;&nbsp; [\faResearchgate](https://www.researchgate.net/profile/J-Baecher?ev=hdr_xprf) &nbsp;&nbsp;&nbsp;&nbsp; [\faOrcid](https://orcid.org/0000-0003-0247-5758) -->

 

# APPOINTMENTS

- **Assistant Professor of Applied Ecology** \hfill Aug 2026 – Present 
  - *Emory University* \| Department of Environmental Science
-  
- **Postdoctoral Fellow in Quantitative Ecology** \hfill July 2025 – July 2026 
  - *Utah State University* \| Department of Rangeland Resources
-  
- **Postdoctoral Research Associate** \hfill Aug 2024 – July 2025 
  - *University of Florida* \| Institute for Food and Agricultural Science
-  
- **Doctoral Preeminence Fellow** \hfill Aug 2019 – Aug 2024 
  - *University of Florida* \| Department of Wildlife Ecology
-  
- **Conservation and Research Technician** \hfill Jan 2018 – July 2019 
  - *Memphis Zoo* \| Department of Conservation and Research
-  
- **NSF REU Research Coordinator** \hfill May 2017 – Aug 2017 
  - *Eastern Kentucky University* \| Department of Biology
-  
- **Graduate Teaching Assistant** \hfill May 2015 – Aug 2017 
  - *Eastern Kentucky University* \| Department of Biology
-  
- **Research Assistant** \hfill Mar 2012 – Mar 2015 
  - *University of Arkansas* \| Department of Biology
-  

# EDUCATION

- **Doctor of Philosophy, Interdisciplinary Ecology** \hfill Aug 2024 
  - *University of Florida*
-  
- **Master of Science, Biology** \hfill Dec 2017 
  - *Eastern Kentucky University*
-  
- **Bachelor of Science, Biology** \hfill May 2014 
  - *University of Arkansas*

# PUBLICATIONS

   **Reverse chronological** • **§Mentee co-author** • **Citations:** ***438*** • **h-index:** ***8***

## Published

2026

17. §Akshay, VA, **JA Baecher**, N Burkett-Cadena, J Thorson, §Y Sánchez, §Y Tavares, §A Bauer, RP Guralnick, LP Campbell. ***2026***. *Linking hosts, landscapes, and climate to advance zoonotic arbovirus forecasting*. ***Scientific Reports*** [***html***](https://www.nature.com/articles/s41598-026-46902-2) [***code***](https://github.com/Campbell-Lab-FMEL/EEEV_forecasting) [***preprint***](https://doi.org/10.32942/X2VM0K)

<!-- -->

16. Oliveira, BF, et al. \[+21 co-authors including **JA Baecher**\]. ***2026***. *Species range shifts often speed ahead of their climatic niches*. ***Proceedings of the National Academy of Sciences*** [***html***](https://doi.org/10.1073/pnas.2515903123) [***preprint***](https://doi.org/10.32942/X2VM0K) [***code***](https://github.com/)
    -    **Press:** [***Nature EcoEvo***](https://drive.google.com/uc?export=download&id=1MhKCoEY3apK7xUfGbH51nua8gCDAoxQ_), [***Conservation Science Partners***](https://csp-inc.org/species-range-shifts-are-more-complex-than-what-the-models-predict/)

2025

15. **Baecher, JA**, §VA Akshay, RP Guralnick, §AM Bauer, §Y Tavares, §Y Sánchez, LP Campbell. ***2025***. *Toward ecological forecasting of West Nile virus in Florida: Insights from two decades of sentinel chicken monitoring*. ***Science of the Total Environment*** *1000:180308* [***html***](https://doi.org/10.1016/j.scitotenv.2024.180308) [***pdf***](https://doi.org/10.1016/j.scitotenv.2024.180308) [***code***](https://github.com/)
    -    **Press:** [***Florida Museum of Natural History***](https://www.floridamuseum.ufl.edu/science/disease-experts-team-up-with-florida-museum-of-natural-history-to-create-a-forecast-for-west-nile-virus/), [***The Microbiologist***](https://www.the-microbiologist.com/news/disease-experts-upgrade-sentinel-chicken-system-to-create-forecast-for-west-nile-virus/6766.article)

2024

14. Evans, LJ, **JA Baecher**, BR Scheffers. ***2024***. *Invasion risk posed by the pet trade*. ***Frontiers in Ecology and the Environment*** [***html***](https://doi.org/10.1002/fee.2825) [***pdf***](https://drive.google.com/uc?export=download&id=1si7rJMDNEE8_ru6V7rErCfMQ0IahGn4P) [***data***](https://doi.org/10.5281/zenodo.13381099)

<!-- -->

13. **Baecher, JA**. ***2024***. *The environmental filtering paradigm and non-filtering community assembly processes*. ***Journal of Biogeography*** *51(11):2109–2116* [***html***](https://doi.org/https://doi.org/10.1111/jbi.14973) [***pdf***](https://drive.google.com/uc?export=download&id=1OqjBtStpDJ3qUTWFgz04Szo5pmbxA_-X)

<!-- -->

12. Klinges, DH, **JA Baecher**, JJ Lembrechts, IMD Maclean, J Lenoir, C Greiser, M Ashcroft, LJ Evans, MR Kearney, J Aalto, LC Barrio, P De Frenne, J Guillemot, K Hylander, T Jucker, M Kopecky, M Luoto, M Macek, I Nijs, J Urban, L van den Brink, P Vangansbeke, J Von Oppen, J Wild, J Boike, R Canessa, M Nosetto, A Rubtsov, J Sallo-Bravo, BR Scheffers. ***2024***. *Proximal microclimate: moving beyond spatiotemporal scale to improve predicted biological responses*. ***Global Ecology and Biogeography*** *33(7):e1–16* [***html***](https://doi.org/https://doi.org/10.1111/geb.13884) [***pdf***](https://drive.google.com/uc?export=download&id=1xq5iYKTIJZYZ7VJPSt6Nk1GyRaRZQDG7) [***code***](https://github.com/ilyamaclean/microclimf)
    -    **Award:** [***Wiley's Top 10 Most Cited***](https://view.email2.wiley.com/?vawpToken=MKYPUOPTL6JEZJTOTOQBYCR3QQ.500023&utm_source=sfmc&utm_medium=email&utm_campaign=B2CJNL_MIS_1693076_Top_Cited_Authors_EM1_T)

<!-- -->

11. Lawlor, JA, L Comte, G Grenouillet, J Lenoir, **JA Baecher**, \[...\], J Sunday. ***2024***. *Mechanisms, detections, and impacts of species redistributions under climate change*. ***Nature Reviews Earth & Environment*** *5:351–368* [***html***](https://www.nature.com/articles/s43017-024-00527-z) [***pdf***](https://drive.google.com/file/d/1Y9WQ5nf7nGaotehWYgnO7NXfUGPRFZrF/view?usp=sharing)
    -    **5-yr Impact Factor =** ***71.5***, **Top 5% Altmetric score**
    -    **Press:** [***list of articles***](https://nature.altmetric.com/details/162629735/news)

<!-- -->

10. Comte, L, R Bertrand, SE Diamond, LT Lancaster, ML Pinsky, BR Scheffers, **JA Baecher**, RMWJ Bandara, I-C Chen, JA Lawlor, NA Moore, BR Oliveira, J Murienne, J Rolland, MA Rubenstein, J Sunday, LM Thompson, F Villalobos, SR Weiskopf, J Lenoir. ***2024***. *Bringing traits back in the equation: A roadmap to understand species redistribution*. ***Global Change Biology*** *30:e17271* [***html***](https://onlinelibrary.wiley.com/doi/epdf/10.1111/gcb.17271) [***pdf***](https://drive.google.com/file/d/1LAGXu1MTVw7XXzsNXgh0JdX6byWyXMCl/view?usp=sharing)

<!-- -->

2023

9.  **Baecher, JA**, SA Johnson, EA Roznik, BR Scheffers. ***2023***. *Experimental evaluation of how invasions and climate change interact to alter the vertical assembly of an amphibian community*. ***Journal of Animal Ecology*** *92(4):875–888* [***html***](https://besjournals.onlinelibrary.wiley.com/doi/10.1111/1365-2656.13899) [***pdf***](https://drive.google.com/file/d/1xpH3gddbvwZFhyESNdwEgNIqN4bKMdB2/view?usp=sharing) [***code***](https://github.com/slamander/treefrog)
    -    **Press:** [***Nature Climate Change***](https://www.nature.com/articles/s41558-023-01654-z.epdf?sharing_token=3XnTm99a17DVcWMhv13iKNRgN0jAjWel9jnR3ZoTv0MMd4D4RpznrkeWRsgunYu7dzCGDugmPYuA8lqoIfQTHmtQ-tp_uQ2CJKXMzH4cKcJkMgfEVbdniiX4uXaKrAHCZLPQwjVg5hZKpGbAUwr8-fFSRsRcdJO9tuJhYTvvjjw%3D), [***Florida Today***](https://www.floridatoday.com/story/news/local/environment/2023/04/03/cuban-tree-frog-hops-to-new-heights-of-ecological-destruction-in-florida/70064687007/), [***Nature Climate Change***](https://secasc.ncsu.edu/2023/03/13/se-casc-researchers-show-how-climate-change-and-invasive-species-threaten-florida-treefrogs/)

2022

8.  Basham, EW, **JA Baecher**, DH Klinges, BR Scheffers. ***2022***. *Vertical stratification patterns of tropical forest vertebrates: A meta-analysis*. ***Biological Reviews*** *98(1):99–114* [***html***](https://doi.org/10.1111/brv.12896) [***pdf***](https://drive.google.com/file/d/1YD5sKz4j02oRsupH4jXtO9l4gnGHdTqD/view?usp=sharing) [***code***](https://github.com/schefferslab/Vert-Strata-Meta)

<!-- -->

7.  Basham, EW, M González-Pinzón, A Romero-Marcucci, N Carl, **JA Baecher**, BR Scheffers. ***2022***. *Large, old trees define the vertical, horizontal, and seasonal distributions of a poison frog*. ***Oecologia*** *199:257–269* [***html***](https://doi.org/10.1007/s00442-022-05108-9) [***pdf***](https://drive.google.com/file/d/1MZieViEYzTJNZoKWg-I4wv0EC8V033yF/view?usp=sharing) [***code***](https://github.com/schefferslab/Basham_et_al_2021)

2020

6.  Swartwout, M, PN Vogrinc, **JA Baecher**, CS Kross, JD Willson. ***2020***. *Prey Size and Growth Rate do not Influence Trophic Morphology of Juvenile Water Snakes* Nerodia sipedon. ***Herpetologica*** *76(1):53–60* [***html***](https://doi.org/10.1655/Herpetologica-D-18-00007)[***pdf***](https://drive.google.com/file/d/1QLdpK0d1k3fEqUHtLrN4H8rV0kjbA-Tc/view?usp=sharing)

<!-- -->

5.  **Baecher, JA**, §E Jones, SC Richter. ***2020***. *Tail bifurcation in two species of* Desmognathus *salamander (Caudata: Plethodontidae) in southeastern Kentucky, USA*. ***Herpetological Bulletin*** *150:29–30* [***html***](https://doi.org/10.33256/hb150.2930) [***pdf***](https://drive.google.com/file/d/1b64jLaFQEWxNdENYW35R3K8eXZbRORKV/view?usp=sharing)

2019

4.  Hinkson, KM, **JA Baecher**, S Poo. ***2019***. *Cryopreservation and hormonal induction of spermic urine in the smooth-sided toad* Rhaebo guttatus. ***Cryobiology*** *89:109–111* [***html***](https://doi.org/10.1016/j.cryobiol.2019.05.007) [***pdf***](https://drive.google.com/file/d/1AyvCIzYfoLL5V0OcxVBo2l8aLJsNoDx9/view?usp=sharing)
    -    **Press:** [***The Commercial Appeal***](https://drive.google.com/file/d/1AyvCIzYfoLL5V0OcxVBo2l8aLJsNoDx9/view?usp=sharing), [***USA Today***](https://drive.google.com/file/d/1AyvCIzYfoLL5V0OcxVBo2l8aLJsNoDx9/view?usp=sharing)

2018

3.  **Baecher, JA**, SC Richter. ***2018***. *Environmental gradients in an old-growth Appalachian Forest predict fine-scale distribution, co-occurrence, and abundance of woodland salamanders*. ***Ecology and Evolution*** *8(24):12940–12952* [***html***](https://doi.org/10.1002/ece3.4736) [***pdf***](https://drive.google.com/file/d/1h3l0vT9pABr5yotXYPAnbmpFwPhFIaTc/view?usp=sharing) [***code***](http://doi.org/10.5281/zenodo.1313705)

2017

2.  **Baecher, JA**, PH Vogrinc, JC Guzy, CS Kross, JD Willson. ***2017***. *Herpetofaunal Communities in Restored and Unrestored Remnant Tallgrass Prairie and Associated Wetlands in Northwest Arkansas*. ***Wetlands*** *38:157–168* [***html***](https://doi.org/10.1007/s13157-017-0966-5) [***pdf***](https://drive.google.com/file/d/1wKPzJ1QXOc-znIXJSQXd9MfpWc5nZ9jU/view?usp=sharing)

2014

1.  **Baecher, JA**, PH Vogrinc, JC Guzy, JC Neal, JD Willson. ***2014***. Lithobates areolatus, *Predation*. ***Herpetological Review*** *45(4):681–682* [***pdf***](https://drive.google.com/file/d/1dZi1LI8vY-JLh46hOCLn4v7oYcfTQeWa/view?usp=sharing)

<!-- -->

## Forthcoming

6.  **Baecher, JA**, D Klinges, LJ Evans, CA Romagosa, R Fletcher, BR Scheffers. ***In revision***. *Isolating what limits the spread of an invasive predator*. ***Journal of Applied Ecology*** [***code***](https://github.com/slamander/pythons) [***preprint***](https://advance.sagepub.com/doi/full/10.22541/au.171248322.23946990/v1)

<!-- -->

5.  Thompson, LM, et al. \[+21 co-authors including **JA Baecher**\]. ***In review***. *The RAD (Resist-Accept-Direct) management of species redistribution under climate change*. ***Frontiers in Ecology and the Environment***

<!-- -->

4.  Klinges, DH, **JA Baecher**, L Soifer, BR Scheffers. ***In review***. *Land modification defines climate change as experienced by life on earth*. ***Science***

<!-- -->

3.  Pittman, C, et al., \[+12 co-authors including **JA Baecher**\]. ***Submitted***. *Connecting living and preserved collections: Standardizing and integrating biological data to cultivate information sharing*. ***Biodiversity Informatics***

<!-- -->

2.  **Baecher, JA**, D Klinges, LJ Evans, BR Scheffers. ***In prep***. *Informing species range-shifts with global climate and habitat connectivity*. Prepared for ***One Earth***

<!-- -->

1.  **Baecher, JA**, E Stuber, TJ Clark-Wolf. ***In prep*** . *Mapping range shifts with a novel integrated spatiotemporal N-mixture Bayesian model*. ***Methods in Ecology and Evolution***

 

# SELECT PRESENTATIONS

   **Reverse chronological** • ***†Student mentee presenter***• **Total presentations:** ***41*** ([***see all***](https://drive.google.com/uc?export=download&id=1IEIzljscjUXfYkjzYkFtWCCD6iwYBgSQ))

12. **Baecher, JA**. ***2026***. *Ecology of rapid change: From behavior to movements to range dynamics*. ***Emory University***, *Atlanta, GA* [***slides***](https://drive.google.com/uc?export=download&id=1vUtZ8b9n1s9lPKrPQhpeAffjUDFqGToZ) [***video***](https://drive.google.com/file/d/1VGHDytkEY54jy6FvXvcOUSdMpeQiE7Pd/view?usp=sharing)

<!-- -->

11. **Baecher, JA**. ***2026***. *Spatial structures for ecological models: A tutorial for point-referenced data*. ***Utah State University*** (hybrid) [***slides***](https://drive.google.com/uc?export=download&id=1vUtZ8b9n1s9lPKrPQhpeAffjUDFqGToZ) [***online resource***](https://www.alexbaecher.com/post/ecological-spatial-structures/)

<!-- -->

10. **Baecher, JA**. ***2025***. *Shifting Frontiers: Understanding species abundance range-shifts for informed conservation strategies*. ***Pacific Flyway Partners*** (hybrid) [***slides***](https://docs.google.com/presentation/d/15n9QGWMfxWnEICz6RBEkCwguPxMMJxI9ji6CscJqy8Q/edit?usp=drive_link)

<!-- -->

9.  **Baecher, JA**. ***2024***. *Influence of Global Change on Species' Behavior, Movement, and Ranges*. ***SNRE Seminar Series***, *Gainesville, FL* [***flyer***](https://drive.google.com/file/d/1g1vZohLll60CgW9KSge9BP3k7pI9ho2V/view?usp=sharing)

<!-- -->

8.  **Baecher, JA**, L Evans, BR Scheffers. ***2024***. *Toward knowledge co-production in invasion modeling: Developing novel interactive expert-elicitation software*. ***Invasive Species Research Institute Conference***, *Gainesville, FL* [***slides***](https://docs.google.com/presentation/d/1OgKFEmrzpVtZ7nYYGqpmuNQz2GtzOZnq/edit?usp=sharing&ouid=118161165194611535602&rtpof=true&sd=true)

<!-- -->

7.  **Baecher, JA**, D Klinges, L Evans, BR Scheffers. ***2023***. *Informing Species Range-shifts with Global Connectivity*. ***Species On The Move***, *Ft Myers, FL* [***slides***](https://www.dropbox.com/scl/fi/sb5mq6qbfa83iq01yinjp/Baecher_et_al_2023_SOTM.pptx?dl=0&rlkey=b4tit2c1et8btsfoqrh2n1tpd)

<!-- -->

6.  **Baecher, JA**, L Evans, BR Scheffers. ***2022***. *Predicting the spread of invasive Burmese pythons in Florida*. ***SERISCC Network Webinar*** (virtual) [***slides***](https://www.dropbox.com/s/0hm8z1x6d2e8l07/Baecher_et_al_2022_IALE.pdf?dl=0) [***video***](https://youtu.be/J-WWcCTeMsU)

<!-- -->

5.  **Baecher, JA**, SB Reichling. ***2018***. *Current status of the Louisiana Pinesnake captive breeding program at the Memphis Zoo*. ***USFWS Working Group Meeting***, *Shreveport, LA*

<!-- -->

4.  **Baecher, JA**, SC Richter. ***2018***. *Natural gradients influence the fine-scale distribution and abundance of salamanders in an old growth Appalachian forest*. ***Joint Meeting of Ichthyology and Herpetology***, *Rochester, NY*

<!-- -->

3.  **Baecher, JA**, P Vongrinc, JD Willson. ***2016***. *Herpetofauna of degraded and restored remnant Tallgrass Prairie*. ***Joint Meeting of Ichthyology and Herpetology***, *New Orleans, LA*

<!-- -->

2.  **Baecher, JA**, DR Leasure, D Lynch, DD Magoulick. ***2013***. *Impact of anthropogenic land use and hydrologic disturbance on crayfish assemblages*. ***Joint Meeting of Aquatic Sciences***. *Portland, OR*

<!-- -->

1.  **Baecher, JA**, J Ortega, SJ Beaupré. ***2013***. *Use of stable isotope analysis for determining the diet breadth of the Western Cottonmouth* (Agkistrodon Piscivorous). ***Society for Freshwater Science Annual Meeting***, *Jacksonville, FL*

<!-- -->

 

# TEACHING

## Instructor of Record

- **University of Florida** \| School of Natural Resources and Environment
  - Environmental Science Lab \hfill Fall 2022 & 2023 

## Teaching Assistant

- **Eastern Kentucky University** \| Department of Biology
  - Human Anatomy Lab \hfill Spring & Fall 2015 – 2017 

## Guest Lectures

- **Utah State University** \| Department of Wildland Resources
  - Landscape Ecology \hfill Fall 2025 
-  
- **University of Florida** \| School of Natural Resources and Environment
  - Community Science for Ecologists \hfill Spring 2025 
-  
- **University of Florida** \| School of Natural Resources and Environment
  - Ecology of Climate Change \hfill Spring 2023 
-  
- **University of Florida** \| School of Natural Resources and Environment
  - Wildlife of Florida \hfill Spring 2022 
-  
- **Eastern Kentucky University** \| Department of Biology
  - Applied Learning in Biology \hfill Spring 2016 
-  
- **Radford University** \| Department of Biology
  - Appalachian Herpetology \hfill Spring 2016 
-  

# STUDENT MENTORSHIP

   **Grad student** (MS, PhD) • Φ **author** • Ψ **conference** • Ω **committee** • → **career prep**

- **Eastern Kentucky University** \hfill 2015 – 2017
  - Alex Gibson, Emily Jones (Φ, Ψ, →MS), Sarah Ashbrook (Ψ), Peyton Blair (Ψ), [Peter Grap](https://www.linkedin.com/in/peter-grap-2238a0158/) (Ψ, →MS, →PhD), Kelley Hoefer (Ψ), Myesha James (Ψ), [Jacob Peters](https://geospatial.yale.edu/profile/jacob-peters) (Ψ, →MS, →PhD), [Julia Portmann](https://geospatial.yale.edu/profile/julia-portmann) (Ψ, →MS, →PhD), [Jocelyn Stalker](https://jocelynstalker.github.io/) (Φ, Ψ, →MS, →PhD)\
     
- **Rhodes College** \hfill 2018 - 2019
  - ([Madeline Estes](https://www.linkedin.com/in/madeline-estes-b65887140/), Ψ, Ω, →MS)\
     
- **University of Florida** \hfill 2019 – 2025
  - Brittany Arthur, Noah Bowen, Sierra Scauzillo (Ψ, Ω), [V Anand Akshay](https://scholar.google.com/citations?user=Kw8YWeMAAAAJ&hl=en) (Φ, Ψ, PhD), [Amy Bauer](https://github.com/amybauer) (Φ, Ψ, PhD), [Yasmin Tavares](https://scholar.google.com/citations?user=bOuCqlYAAAAJ&hl=en) (MS, Φ, Ψ, →PhD), [Yesenia Sanchez](https://www.linkedin.com/in/yesenia-s%C3%A1nchez-aa1188273/) (Φ, Ψ)\
     
- **Utah State University** \hfill 2025 – Present
  - Jack Rasmussen (PhD), [Mallory Sandoval-Lambert](https://scholar.google.com/citations?user=W_sZJ2EAAAAJ&hl=en) (MS), [Adrien Bastidas](https://www.linkedin.com/in/adrien-bastidas-581b4127a/) (PhD)

 

# WORKSHOPS

   **Reverse chronological**

1.  Fletcher, RJ, **JA Baecher**, M Lezzi, A Marx. ***2023***. *Application of connectivity modeling to identify and predict movement and range redistribution*. ***Species on the Move***, *Bonita Springs, FL* [***slides***](https://docs.google.com/presentation/d/1jsUhEewx3OWyfGIZvX3fCAWCGYeYyffM/edit?usp=sharing&ouid=118161165194611535602&rtpof=true&sd=true)

2.  **Baecher, JA**, L Evans, BR Scheffers. ***2022***. *Expert elicitation of key invasive species habitat requirements, prevalence, and life history attributes*. ***Austin Cary Forest***, *Gainesville, FL* [***slides***](https://drive.google.com/file/d/1JXUdEwaIEClBa5bnz4BMLdO-aUTqmZ5i/view?usp=sharing)

3.  Fletcher, RJ, **JA Baecher**, M Lezzi, A Marx. ***2022***. *Predicting the spread of invasive Burmese pythons in Florida*. ***Annual Meeting of the International Association of Landscape Ecologists - North America*** (virtual) [***video***](https://www.youtube.com/watch?v=6A4WQrwyIHI&ab_channel=RobertFletcher)

 

# EDUCATIONAL RESOURCES

  **Publically available at Questlab [website](https://questlab.dev)**

-    ***Mixed Effects Models in R*** [\faGlobe](https://questlab.dev/post/mixed-effects-models-in-r/) \hfill 2025 
  -    Open source and continuously updated *Rshiny* tool
-    ***Spatial Structures in Ecological Models*** \| Tutorial [\faGlobe](https://questlab.dev/post/ecological-spatial-structures/) \hfill 2025 
  -    Comprehensive tutorial for spatially-explicit modeling with descriptions
-    ***Intro to R Coding:***
  -    Part 1: the basics (e.g., syntax, operators, functions) [\faGlobe](https://questlab.dev/post/r-intro/) \hfill 2022 
  -    Part 2: advanced techniques (e.g., data wrangling, visualization, for loops) [\faGlobe](https://questlab.dev/post/r-intro-pt2/) \hfill 2022 
-    ***Machine learning the “Tidy” way*** [\faGlobe](https://questlab.dev/post/ml-tidymodels/) \hfill 2021 
  -    Tutorial for developing AI/ML models with *tidymodels* and *carot*
-    ***Linear regression with gradient decent*** [\faGlobe](https://questlab.dev/post/gradient-descent/) \hfill 2021 
  -    Tutorial for developing custom linear models with maximum likelihood and gradient descent
-    ***Mapping ecological flow in R*** [\faGlobe](https://questlab.dev/post/connectivity-script/) \hfill 2020 
  -    Tutorial for connectivity modeling with randomized shortest paths, least cost, and circuit theory
-    ***Empirical Dynamic Models for forecasting*** [\faGlobe](https://questlab.dev/post/edm-script/) \hfill 2016 
  -    Tutorial for forecasting non-linear timeseries data

# FUNDING

  **Reverse chronological**

```{r funding-table}
funding <- tribble(
  ~Amount, ~Organization, ~Fund, ~Year,
  "$680,000", "Emory University", "Lab Start-up Fund", "2026",
  "$130,000", "US Geological Survey", "Climate Adaptation Postdoctoral Fellows", "2025",
  "$195,000", "University of Florida", "Doctoral Preeminence Fellowship Award", "2019",
  "$4,000", "University of Florida", "Robin Nadeau Graduate Research Award", "2019",
  "$5,000", "National Science Foundation", "Research Experience for Undergraduates", "2013",
  "$5,000", "National Science Foundation", "Research Experience for Undergraduates", "2012",
  "$1,100", "Eastern Kentucky University", "Marcia Athey Research Grant", "2016",
  "$800", "Eastern Kentucky University", "Dept. Biology Travel Grant", "2017",
  "$700", "Society Freshwater Science", "Instars Fellowship", "2014",
  "$700", "Society Freshwater Science", "Instars Fellowship", "2013",
  "$500", "Soc. Study Amphibians & Reptiles", "Grant-in-Aid of Herpetological Research", "2016",
  "$500", "Eastern Kentucky University", "Dep. Biological Sciences Travel Stipend", "2016",
  "$500", "Eastern Kentucky University", "Div. Natural Areas Research Grant", "2016",
  "$500", "University of Arkansas", "Dept. Biology Travel Grant", "2013"
)

kable(funding, 
      format = "latex",
      booktabs = TRUE,
      linesep = "",
      align = c("r", "l", "l", "r"),
      col.names = c("Amount", "Organization", "Fund Name", "Year")) %>%
  kable_styling(latex_options = c("scale_down", "HOLD_position"),
                font_size = 11)
```

   **Total Funding:** ***\$1,024,000*** -  

# PROFESSIONAL ACTIVITIES

- **Peer Review** (Total Requested: ***38*** \| Total Reviewed: ***25***)
  - *Biological Conservation, Conservation Science and Practice, Ecography, Ecology, Ecology & Evolution, Frontiers in Ecology and Environment, Herpetological Bulletin, Herpetological Conservation Biology, Herpetological Review, Hydrobiologia, Natural Areas Journal, Nature Ecology & Evolution, Oecologia, Scientific Reports, Wildlife Society Bulletin*
  -  
- **Professional Membership**
  - ***Ecological Society of America*** \hfill 2019 – Present 
  - ***International Association of Landscape Ecologists – North America*** \hfill 2020 – Present 
  - ***USGS Climate Adaptation Science Center***
    - Doctoral Research Fellow \| Southeastern CASC \hfill 2019 – 2024 
    - Postdoctoral Research Fellow \| Southwestern CASC \hfill 2025 – 2026 
  - ***American Society of Ichthyologists and Herpetologists*** \hfill 2014 – 2021 
  - ***Southeastern Partners in Amphibian and Reptile Conservation*** \hfill 2014 – 2021 
  - ***Society for the Study of Amphibians and Reptiles*** \hfill 2014 - 2021 
  - ***Society of Freshwater Ecology*** \hfill 2012 – 2014 
    - Instars Fellow (2012 & 2013)
  -  
- **Working Groups**
  - [***BioShifts***](https://www.fondationbiodiversite.fr/en/the-frb-in-action/programs-and-projects/le-cesab/bioshifts/) (Member, Group Leader) \hfill 2022 – Present 
  - ***DirtClim*** (Core team) \hfill 2019 – Present 
  - [***National Synthesis for Ecological Analysis and Synthesis***](https://www.nceas.ucsb.edu/) (Member) \hfill 2025 – Present 
  - [***ResearchR***](https://researchr.us/) (President of Scientific Integrity) \hfill 2025 – Present 
  - [***ZooMu Network***](https://zoomunetwork.org/) (Member) \hfill 2024 – Present 
  - ***USGS CAP Fellows National Synthesis Effort*** (Postdoctoral Fellow) \hfill 2025 – Present 
  - [***Southeast RISCC***](https://southeastriscc.org/) (Member) \hfill 2022 – Present 

 

# NON-FUNDED PROPOSALS

  **Proposals available upon request**

```{r non_funding-table}
non_funding <- tribble(
  ~Amount, ~Organization, ~Decision, ~Year,
  "$235,000", "NSF: Postdoctoral Research Fellowship – Biology", "Meritorious", "2024",
  "$235,000", "NSF: Postdoctoral Research Fellowship – Biology", "Highly Meritorious", "2023",
  "$25,000", "The Wilderness Society: Gloria Barron Fellowship", "Declined", "2021",
  "$3,500", "American Museum Natural History: Roosevelt Memorial Grant", "Declined", "2020",
  "$4,000", "Foundation for the Conservation of Salamanders", "Declined", "2020",
  "$60,000", "Disney Conservation Grant", "Inquiry declined", "2019",
  "$10,000", "National Park Service: Tanner Memorial Fellowship", "Inquiry declined", "2019"
)

kable(non_funding, 
      format = "latex",
      booktabs = TRUE,
      linesep = "",
      align = c("r", "l", "l", "r"),
      col.names = c("Amount", "Organization", "Decision", "Year")) %>%
  kable_styling(latex_options = c("scale_down", "HOLD_position"),
                font_size = 11)
```

**Total Non-funded:** ***\$572,500***

 

# REFERENCES

   **Since PhD (2019 - Present)**

- ***Robert Guralnick, PhD*** \hfill 2019 – Present 

  - **Curator of Biodiversity Informatics** \| *Florida Natural History Museum*
    -     \faPhone: (352) 273-1980 • \faEnvelope: [***robgur\@gmail.com***](mailto:robgur@gmail.com) • [***Guralnick lab***](mailto:robgur@gmail.com)\

-  

- ***Brett Scheffers, PhD*** \hfill 2019 – Present 

  - **Associate Professor** \| *Dept of Wildlife Ecology, University of Florida*
  - **Director** \| *Florida Climate Institute, University of Florida*
    -     \faPhone: (352) 846-0570 • \faEnvelope: [***brett.scheffers\@ufl.edu***](mailto:brett.scheffers@ufl.edu) • \faGlobe: [***Scheffers lab***](http://schefferslab.com/)\

-  

- ***Lindsay Campbell, PhD*** \hfill 2024 – Present 

  - **Assistant Professor** \| *Florida Medical Entomology Lab, University of Florida*
    -     \faPhone: (772) 226-6666 • \faEnvelope: [***lcampbell2\@ufl.edu***](mailto:lcampbell2@ufl.edu) • \faGlobe: [***Campbell lab***](http://lcampbelllab.wixsite.com/campbell-lab)\

-  

- ***TJ Clark-Wolf, PhD*** \hfill 2025 – Present 

  - **Assistant Professor** \| *Dept of Wildland Resources & Ecology Center, Utah State University*
    -     \faPhoneSlash: email preferred • \faEnvelope: [***t.j.clark-wolf\@usu.edu***](mailto:t.j.clark-wolf@usu.edu) • \faGlobe: [***Clark-Wolf lab***](https://www.clarkwolflab.com/)\

-  

- ***Erica Stuber, PhD*** \hfill 2025 – Present 

  - **Assistant Unit Leader** \| *USGS UT Cooperative Fish and Wildlife Research Unit*
  - **Assistant Professor** \| *Dept of Wildland Resources & Ecology Center*
    -     \faPhoneSlash: email preferred • \faEnvelope: [***erica.stuber\@usu.edu***](mailto:erica.stuber@usu.edu) • \faGlobe: [***Stuber lab***](https://efstuber.wixsite.com/xscaleeco)